---
title: "ASA SMI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
```

# Packages

Some packages that we need to load for analyses like `tidyverse` and `spatialTIME` available from CRAN.

```{r packages, message=FALSE}
#make vector of packages to install and load
packages = c("tidyverse", "spatialTIME", "here")
#check if packages are installed and if not, install
install.packages(setdiff(packages, rownames(installed.packages())))
t = lapply(packages, require, character.only = T)
```

# Importing Data

The clinical data, per-cell spatial data, and sample summary data are in the `data/` subdirectory

```{r data_import}
#read clinical csv
clinical_data <- read.csv(here::here("data/deidentified_clinical.csv"))
head(clinical_data)
#read summary csv
sample_summary <- read.csv(here::here("data/deidentified_summary.csv"), sep="")
head(sample_summary)
#read spatial list
load(here::here("data/deidentified_example.RData"))
names(example_spatial_small)
```

# Create Multiplex ImmunoFlourescent (mif) Object 

spatialTIME functions use a custom `mif` object which can be created 
using the `create_mif()` function. 
The `mif` object has 6 slots storing the: 

* Clinical data which must contain:
  - a column whose column name matches one in the sample dataset.
* Sample summary data including counts/percentages of each positive cells
for each single or combination of markers and total number of cells for 
each core. In order to use `create_mif()` function this table must contain:
  - a column whose column name matches one in the clinical dataset
  - a column whose column name matches one in the spatial list.
* Spatial list (1 per each core):
  - This object should be a list object where each element of the list
  corresponds to a core and each element should be a  $n\times p$ 
  dataframe ($n$ = number of cells)  containing:
    1. a column name that matches a column name for the sample file 
    (for merging and potential downstream analysis linking to clinical
    variables),
    2. `XMin`, `XMax`, `YMin`, and `YMax` which defines the area that a
    cell occupies which is eventually used to assign a location for each
    cell with the x-position being the mean of the `XMin` and `XMax` and
    y-position being the mean of the `YMin` and `YMax`
    3. a set of columns that indicate whether a cell is positive to one
    or multiple markers.
* patient id:
  - a column name used to merge clinical and summary data
* sample_id:
  - a column name used to merge the spatial and summary data
* derived:
  - where all of the plots and spatial clustering measures are stored

# Creating MIF object

```{r ceate_mif_object}
# Make sure the variable types are the same for deidentified_id and 
# deidentified_sample in their corresponding datasets
x <- spatialTIME::create_mif(clinical_data = clinical_data %>%
                               mutate(deidentified_id = as.character(deidentified_id)),
                             sample_data = sample_summary %>%
                               mutate(deidentified_id = as.character(deidentified_id)),
                             spatial_list = example_spatial_small,
                             patient_id = "deidentified_id", sample_id = "deidentified_sample")

x #prints a summary of how many patients, samples, and spatial files are present

```

# Visualizing TMAs

A plot of for each sample in the spatial list is created and assigned to the `derived`
slot in the mif object. If an argument to `filename` is provided then a PDF with one 
figure per page is created. 

The order of the phenotypes provided matters. When phenotypes are derived by multiple markers, 
the last provided phenotype will overlap previous phenotypes - so put the marker combinations 
before the individual markers. 

```{r mif_plotting, warning = FALSE, message = FALSE}
names_one <- c("CD3..CD8.","CD3..FOXP3.","CD3..Opal.570..Positive",
               "CD8..Opal.520..Positive","FOXP3..Opal.620..Positive", 
               "PDL1..Opal.540..Positive", "PD1..Opal.650..Positive")

names_two <- c("CD3..Opal.570..Positive","CD8..Opal.520..Positive",
               "FOXP3..Opal.620..Positive","PDL1..Opal.540..Positive",
               "PD1..Opal.650..Positive","CD3..CD8.","CD3..FOXP3.")

#add an element in the `derived` object position
x <- spatialTIME::plot_immunoflo(x, plot_title = "deidentified_sample",  
                                 mnames = names_one, cell_type = "Classifier.Label")
x[["derived"]][["spatial_plots"]][[4]] +
  scale_color_manual(breaks = names_two, 
                     values = viridis::turbo(7),
                     labels = names_two %>%
                       gsub("..Opal.*", "+", .) %>% 
                       gsub("\\.\\.", "+", .) %>% 
                       gsub("\\.", "+", .))

# name order matters 
x <- spatialTIME::plot_immunoflo(x, plot_title = "deidentified_sample",  
                                 mnames = names_two, cell_type = "Classifier.Label")
x[["derived"]][["spatial_plots"]][[4]] +
  scale_color_manual(breaks = names_two, 
                     values = viridis::turbo(7),
                     labels = names_two %>%
                       gsub("..Opal.*", "+", .) %>% 
                       gsub("\\.\\.", "+", .) %>% 
                       gsub("\\.", "+", .))
  
```

# Complete spatial randomness

Complete spatial randomness (CSR) is the assumption that the cells are randomly distributed across the region,
with no evidence of clustering or repulsion, and that the cell intensity is constant across the entire region.
When tissue cores are collected, rips or tears can create regions where it appears there are no cells (when
this is not the case). This violates the CSR assumption and therefore theoretical estimates of CSR may
not be accurate.
Damage can occur to tissue cores due to how they are collected. This damage can lead to rips and tears in the
cores which results in regions where it appears that cells are not located and is not actually the case. Due
to these violations of the CSR assumption, the theoretical estimate for CSR may not be accurate.

# Univariate analysis

The spatialTIME package can compute 3 count-based spatial statistics: Ripley's K, 
Besag's L and Marcon's M. 
Ripley's K measures the average number of cells within a specified radius of a cell. 
Edge corrections for Ripley's K are used to account for points outside of the observed region and assume that the 
distribution of cells is the same inside and outside of the observed region. 
There are 3 main edge corrections available in spatialTIME - isotropic,
translational or border. We recommend using either isotropic or translational edge corrections 
when there are a small number of cells. 
The package can also compute a distance-based measure: G, which is the proportion of cells whose distance
to its nearest neighbor is less than a specified amount (`r`). 
Edge corrections for G apply different cell inclusion criteria and are either reduced sample and Hasnisch. 

## Univariate - count

The `ripleys_k` function reports a permuted and theoretical estimate of CSR, the observed value for the 
specified statistic, and the full permutation distribution of the statistic if `keep_perm_dis = TRUE`.

In this example, the number of permutations is 10, but this should be increased to closer to 100 
for a more reliable estimate when doing a real analysis.

```{r ripleys, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = 'center'}
x <- ripleys_k(mif = x, mnames = names_two, 
               num_permutations = 10, method = "K",
               edge_correction = 'translation', r = seq(0,100,10),
               keep_perm_dis = FALSE, workers = 1)

tail(x$derived$univariate_Count)

values = viridis::turbo(length(unique(x$derived$univariate_Count$deidentified_sample)))
names(values) = unique(x$derived$univariate_Count$deidentified_sample)

x$derived$univariate_Count  %>%
  filter(Marker != 'PD1..Opal.650..Positive') %>%
  ggplot(aes(x = r, y = `Degree of Clustering Permutation`)) +
  geom_line(aes(color = deidentified_sample), show.legend = FALSE) +
  facet_wrap(Marker~., scales = 'free') + 
  theme_bw() + 
  scale_color_manual(values = values)
  
```

Positive values for degree of cluster when using `method = 'K'` and `method = 'L'` indicates
evidence of spatial clustering, while negative values correspond to spatial regularity. 

On the other hand, if using `method = 'M'` then values > 1 correspond to clustering
and values < 1  correspond to regularity. 
These values can also be interpreted as the the percent difference from spatial clustering,
for example if $M=0.5$ that means there is 50% less spatial clustering than expected under CSR.

## Univariate - nearest neighbor

```{r NN, warning = FALSE, fig.width = 10, fig.height = 6, fig.align = 'center'}

x <- NN_G(mif = x, mnames = names_two, num_permutations = 10,
                edge_correction = 'rs', r = seq(0,100,10),
                keep_perm_dis = FALSE, workers = 1)

tail(x$derived$univariate_NN)

x$derived$univariate_NN  %>%
  filter(Marker != 'PDL1..Opal.540..Positive') %>%
  ggplot(aes(x = r, y = `Degree of Clustering Permutation`)) +
  geom_line(aes(color = deidentified_sample)) +
  facet_wrap(Marker~., scales = 'free') + theme_bw() + 
  scale_color_manual(values = values)
  
```

The interpretation of the degree of clustering for $G$ that values > 0 
indicate spatial clustering of the cell types of interest, while values < 0
indicate dispersion of these cells.

# Bivariate analysis 

## Bivariate - count 

In bivariate analysis, we are interested in how many cells of Type 1 (Counted) 
are clustered in proximity to Type 2 (Anchor). 
Essentially the radius is centered around a cell of Type 2 and then the cells 
of Type 1 are counted.

```{r, fig.width = 10, fig.height = 6, fig.align = 'center'}
x <- bi_ripleys_k(mif = x, mnames = c("CD3..CD8.", "CD3..FOXP3."), 
                  num_permutations = 10, method = "K",
                  edge_correction = 'translation', r = seq(0,100,10),
                  keep_perm_dis = FALSE, workers = 1, exhaustive = TRUE)

head(x$derived$bivariate_Count)

x$derived$bivariate_Count  %>%
  filter(anchor == 'CD3..FOXP3.') %>%
  ggplot(aes(x = r, y = `Degree of Clustering Permutation`)) +
  geom_line(aes(color = deidentified_sample), show.legend = TRUE) +
  theme_bw() + scale_color_manual(values = values)
```

## Bivariate - nearest neighbor 

```{r fig.width = 10, fig.height = 6, fig.align = 'center'}
x <- bi_NN_G(mif = x, mnames = c("CD3..CD8.", "CD3..FOXP3."), num_permutations = 10,
               edge_correction = 'rs', r = seq(0,100,10),
               keep_perm_dis = FALSE, workers = 1, overwrite = TRUE)

head(x$derived$bivariate_NN)

x$derived$bivariate_NN  %>%
  filter(anchor == 'CD3..FOXP3.') %>%
  ggplot(aes(x = r, y = `Degree of Clustering Permutation`)) +
  geom_line(aes(color = deidentified_sample), show.legend = TRUE) +
  theme_bw() +  scale_color_manual(values = values)
```

## Additional analyses

```{r, fig.width = 10, fig.height = 6, fig.align = 'center'}
x <- ripleys_k(mif = x, mnames = names_two, num_permutations = 10,
               edge_correction = 'translation', r = c(0,50),
               keep_perm_dis = FALSE, workers = 1, overwrite = TRUE)

inner_join(x$clinical, x$derived$univariate_Count) %>%
  filter(grepl('CD3..O|CD8..O', Marker)) %>% 
  ggplot(aes(x = status, y = `Degree of Clustering Permutation`)) + 
  geom_point(aes(color = deidentified_sample), show.legend = FALSE) + 
  facet_wrap(Marker ~., scales = 'free') 

```
