---
title: "ASA SMI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create Multiplex ImmunoFlourescent (mif) Object 

spatialTIME functions use a custom `mif` object which can be created 
using the `create_mif()` function. 
The `mif` object has 6 slots storing the: 

* Clinical data which must contain:
  - a column whose column name matches one in the sample dataset.
* Sample summary data including counts/percentages of each positive cells
for each single or combination of markers and total number of cells for 
each core. In order to use `create_mif()` function this table must contain:
  - a column whose column name matches one in the clinical dataset
  - a column whose column name matches one in the spatial list.
* Spatial list (1 per each core):
  - This object should be a list object where each element of the list
  corresponds to a core and each element should be a  $n\times p$ 
  dataframe ($n$ = number of cells)  containing:
    1. a column name that matches a column name for the sample file 
    (for merging and potential downstream analysis linking to clinical
    variables),
    2. `XMin`, `XMax`, `YMin`, and `YMax` which defines the area that a
    cell occupies which is eventually used to assign a location for each
    cell with the x-position being the mean of the `XMin` and `XMax` and
    y-position being the mean of the `YMin` and `YMax`
    3. a set of columns that indicate whether a cell is positive to one
    or multiple markers.
* patient id:
  - a column name used to merge clinical and summary data
* sample_id:
  - a column name used to merge the spatial and summary data
* derived:
  - where all of the plots and spatial clustering measures are stored


```{r data_import}

```